Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\

### 2025-03-07 main-disk-model.tex
main-disk-model.tex cleaned until II.C

        %Scheduling scientific workflows is important. (\skug{can't imagine a good first sentence})
        Scientific workflows are often represented as directed acyclic graphs (DAGs),
        where vertices correspond to tasks and edges represent the dependencies between them.
        Typically, each task requires a
        certain amount of memory to be executed and needs to communicate data to its successor tasks.
        The goal is generally to execute the workflow  as fast as possible (i.e., to minimize its makespan),
        while satisfying the memory constraints.
        Hence, we investigate the memory-aware scheduling of DAG-shaped workflows on
        heterogeneous platforms, where each processor can have a different speed and a different memory size.
        We propose a variant of HEFT (Heterogeneous Earliest Finish Time) that (in contrast to the original) accounts for memory and
        includes eviction strategies for cases when it might be beneficial to remove some data from memory
	in order to have enough memory to execute other tasks.
%
	Furthermore, while HEFT assumes perfect knowledge of the execution time and memory usage
	of each task, the actual values might differ upon execution. Thus, we propose an adaptive
	scheduling strategy, where a schedule is recomputed when there has been a significant variation in terms
	of execution time or memory.
%
	The scheduler has been closely integrated with a runtime system, allowing us to perform a thorough
	experimental evaluation on real-world workflows. The runtime system warns the scheduler when
	the task parameters have changed, and a schedule can be recomputed on the fly. The memory-aware
	strategy allows us to schedule task graphs that would run out of memory with a state-of-the-art
	scheduler, and the adaptive setting allows us to significantly reduce the makespan.

%                when tentatively assigning tasks, in order to
%        Its first step is to compute the weights of the tasks.
%        We suggest three variants: bottom levels as weights, bottom levels with impact of incoming edge weight,
%        and weights along the optimal memory traversal.
%        In the second step, we try assigning each task to each processors and execute the assignment that
%        is feasible with regard to memory size and gives the earliest finishing time to the task.
%        Sometimes, data corresponding to edge weights that is stored in the memory needs to be evicted in order to
%        assign a task to a processor.
%        We suggest two eviction strategies - largest files first and smallest files first.

%        Our experimental evaluation on real-world workflows and simulated (\skug{generated? They are generated from real-world wfs})
%        ones with real task and edge weights with up to 30,000 tasks shows that
%        respecting memory constraints only costs $11\%$ of runtime in comparison to a non memory-aware baseline.
%        Calculating task weights with the impact of memory gives a $x\%$ better makespans on a normal and $y\%$ better makespans
%        on small one.
%        Calculating task weights along the optimal memory traversal gives on average $z\%$ worse makespans, but improves
%        average memory utlization by $t\%$.

@@

\section{Introduction} %: \skug{Full: 0, Polished: 0}}

%    \skug{Fullness score (how much text is available): 0-none to 5 - everything.
%
%    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
%    }
%    TODO: Insert abstract

    %%% CONTEXT %%%
    The analysis of massive datasets, originating from fields such as genomics, 
    remote sensing, or biomedical imaging -- to name just a few -- has become ubiquitous in science;
    this often takes the form of workflows, \iec separate software components chained together
    in some kind of complex pipeline~\cite{DBLP:journals/dbsk/LeserHDEGHKKKKK21}.
    These workflows are usually represented as directed acyclic graphs (DAGs).
    The DAG vertices represent the software components (or, more generally, the workflow \emph{tasks}),
    while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liu2018survey}.
    Large workflows with resource-intensive tasks can easily exceed the capabilities of a 
    single computer and are therefore executed on a parallel or distributed platform.
    An efficient execution of the workflows on such platforms requires mapping tasks
    to specific processors; to increase utilization by reusing finished processors,
    one also needs a task schedule (\iec a valid execution order that respects the dependencies)
    and possibly also starting times for the tasks.

    %%% MOTIVATION %%%
    Modern computing platforms are often heterogeneous, meaning they feature varying CPU speeds
    and memory sizes. In general, having different memory sizes per CPUs makes it more challenging to compute
    a schedule that respects all memory constraints -- meaning that no task is executed on a 
    processor with less memory than needed for the task. This is, however, very important to
    avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
    Hence, building on previous related %\hmey{You may want to add other refs not from us} 
    work~\cite{gou2020partitioning,He21,DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
    formulation that takes memory sizes as explicit constraints into account. Its objective is
    the very common \emph{makespan}~\cite{liu2018survey}, 
    which acts as proxy for the total execution time of a workflow.
However, to the best of our knowledge,
    the only memory-aware heuristics that would account for memory constraints are  partitioning
    the DAG and not reusing processors once they have processed a part of the graph, leading
    to high values of makespan compared to a finer grain solution that reuses processors.

    While previous work with memory constraints has been focusing on partitioning the graph, 
    and not reusing processors during execution, 
    a seminal list scheduling heuristic for workflows on heterogeneous platforms, without accounting
    for the memory constraint, is HEFT 
    (heterogeneous earliest finish time)~\cite{topcuoglu2002performance}.
    It has two phases: (i) each task is assigned a priority; and (ii) the tasks in a priority-ordered list are assigned
    to processors, where the ``ready'' task with the highest priority is scheduled next on the processor
    where it would complete its execution first. 
    HEFT has been extended (e.g., by Shi and Dongarra~\cite{SHI2006665}) and adjusted 
    for a variety of different scheduling problem formulations. 
    Yet, none of them adhere to memory constraints as we propose -- see discussion of related work
    in Section~\ref{sec:related-work}. 
%    (\skug{check in related work if true!}).
%        \skug{Note: 2 papers that deal with memory sizes, but model is very different!}

    Another limitation in practice of HEFT (and many other scheduling strategies) is their 
    assumption that the task running times provided to them are accurate. In practice, this is 
    not the case and deviations from user estimates or historical measurements are 
    very common~\cite{hirales2012multiple}. As a consequence, one should adapt the schedule when \emph{major}
    deviations occur. However, the original list-based schedulers, such as HEFT, are only defined
    in a static setting with accurate task parameters. 
%    
%    List-based schedulers such as HEFT are, however, not designed for 
%    such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement? Please add ref (if any)}
%    \AB{Actually, we adapt HEFT as well, I'll reformulate...}
%%    and would compute a completely new schedule from scratch.

    %%% CONTRIBUTION %%%
    %\paragraph*{Contribution} 
    The main contributions of this paper are both algorithmic and experimental:
    \begin{itemize}
\item We formalize the problem with memory constraints, where communication buffers
are used to evict data from memory if it will be later used by another processor. 
\item We design three HEFT-based heuristics
    that adhere to memory size constraints: \heftbl, \heftblc, and \heftmm
    (M behind HEFT for \underline{m}emory, BL for \underline{b}ottom \underline{l}evel,
    BLC for \underline{b}ottom \underline{l}evel with \underline{c}ommunication, 
    and MM for \underline{m}inimum \underline{m}emory traversal).
    The difference between the new heuristics is the way they prioritize tasks (for processor assignment).

    \item We implement a runtime system able to provide some feedback to the scheduler
    when task requirements (in terms of execution time and memory) differ from the initial predictions, 
    and we recompute a schedule, based on the reported deviations.

    \item We perform extensive simulations, first in the static case by comparing the schedules produced 
    by these heuristics with the classical HEFT as baseline, which however does not take memory sizes into account; 
    while HEFT returns invalid schedules that exceed the processor memories and cannot execute correctly,
    the new heuristics are able to successfully schedule large workflows, with reasonable makespans.

    \item In the dynamic setting, we use a runtime system that allows us to simulate workflow executions,
    introducing deviations in running times and memory requirements of tasks that are communicated
    back to the scheduler; the scheduler can then recompute a schedule. Without these recomputations,
    most schedules become invalid after deviations, since the memory constraint is exceeded 
    for most workflows, hence demonstrating the necessity of a dynamic adjustment of the schedule.

%    \hmey{Need to clarify: why is this SotA?}

  %  \begin{itemize}
%      \item Static: we find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.
%      \item Adaptive: runtime system built, simulates workflow executions and deviations in running times and mem requirements of tasks
%      \item Answering requests of the runtime system for adaptation, the scheduler computes an improved schedule based on the reported deviations.
    \end{itemize}

We first review related work in Section~\ref{sec:related-work}. Then, we formalize the model in Section~\ref{sec:model}
and the algorithms in Section~\ref{sec:heuristics}. The adaptation of the heuristics in a dynamic setting is discussed in Section~\ref{sec:dyn}, and experiment results are presented in Section~\ref{sec:expe}. Finally, we conclude
and provide future working directions in Section~\ref{sec:conc}.

\section{Related work} %: \skug{Full: 5, Polished: 4}}
    \label{sec:related-work}

   First, we focus on HEFT-like scheduling heuristics from the literature  that do not necessarily  
   consider memory constraints. Then, we discuss memory-aware scheduling algorithms. 
   Finally, we move to related work on dynamic or adaptive algorithms. 
 %   \hmey{Question: Does it make sense to place related work before the model (which is more common)?}

%    We discuss relevant scheduling approaches that reuse processors or respect the memory requirements of the processors.
%
%    \subsection{Early list schedulers with unlimited processors}
%    An entire cluster of works on list schedulers has been carried out as early as the 90s.
%    They all assume a DAG-shaped workflow with makespan weights on tasks, and an unlimited amount of homogeneous processors
%    with the speed of 1~\cite{benoit2013survey}.\hmey{needs refs or at least a survey pointer}
%    \skug{actually, Anne cited them, so citing her :-)}
%
%    The \textit{task duplication}-based approaches exploit that sometimes running a task twice on different machines can
%    help reduce the makespan by saving communication costs.
%    The two categories are scheduling with partial duplication~(SPD), and with full duplication~(SFD).
%    For a join task (a task whose incoming degree is larger than its outgoing degree), SPD finds a critical immediate
%    parent (the one that gives the largest start time to the join task) and duplicates only it.
%    SFD duplicates all parents of a join node.
%    The algorithm by~\cite{dfrn1997} duplicates first (creates copies of all parent tasks) and then reduces (removes) the ones that can be removed without harming the makespan.
%    The critical path fast duplication algorithm CPFD~\cite{5727760} classifies tasks into three categories: critical
%    path task, in-branch task, or out-branch task.
%    It schedules critical path tasks first, then in-branch tasks.
%    \hmey{I'm missing the connection to our paper. Or vice versa, how is our contribution
%    connected to these works? (Einordnung in den Forschungskontext) For example, do we do similar things? Are there interesting limitations we overcome?}
%
%    Linear clustering~\cite{KWOK1999381} acts on critical paths in the workflow.
%    It assigns the current critical path to one processor, removes all these tasks from the workflow, recomputes the critical
%    path and repeats the procedure.
%    Heaviest node first~\cite{SHIRAZI1990222} assigns the tasks level by level;
%    in each level, it schedules the one with largest computation time first.

    \subsection{Static list schedulers, especially HEFT-based algorithms}

    Introduced in 2002, HEFT~\cite{topcuoglu2002performance} is a list-based heuristic.
    It and all its successors consist of two phases: task prioritization/ordering and task assignment.
    In the first phase, the algorithms compute bottom levels of the tasks based on some priorities (create the list),
    and then schedule tasks in the order of these priorities.
    The modifications of HEFT revolve around the way the priorities of the tasks are computed and the logic of the processor assignment.
    All such algorithms assume a heterogeneous execution environment.

    Hence, during the task prioritization phase in~\cite{sulaiman2021hybrid},  the standard deviation of the computation cost
    (between processors) is computed, and added to the mean value to account for the differences between processor speeds.
    In the processor choice phase, the entry task and the longest predecessor tasks are duplicated 
    during idle times on the processor.

%    Ref.~\cite{alebrahim2017task} computes the bottom level based on the difference of execution times on
%    the fastest and the slowest processors, divided by the speed ratio of these two processors.
%    When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
%    They choose the processor with the lowest execution time and cross over to other processors sometimes.
%    They build upon~\cite{shetti2013optimization}.\hmey{Last sentence ``hangs in the air''. Either drop or connect it properly.}

    PEFT (Predict earliest finish time)~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
    Cost Table (OCT).
    The OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task if this
    processor is chosen for this task.
    Ranking is based on OCT values.
    The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task.

    The HSIP (Heterogeneous Scheduling with Improved task Priorities)~\cite{wang2016hsip} has an improved first step in
    comparison to HEFT.
    It combines the standard deviation with the communication cost weight on the tasks.
    In the second stage, the algorithm duplicates the entry task if there is a need for it.

    The TSHCS (Task Scheduling for Heterogeneous Computing Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
    by adding randomized decisions to the second phase.
    The decision is whether the task be assigned to the processor with the lowest execution time or to the processor that
    produces the lowest finish time.

    The SDC algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to task’s
    average execution cost in its weight.
    The selected task is then assigned to a processor which minimizes its Adjusted Earliest Finish Time (AEFT) that
    additionally notes how large the communication between current node and its children will be on the
    average provided that it is scheduled on the current processor.

    HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement learning techniques~\cite{yano2022cqga}.

    \subsection{Memory-aware scheduling algorithms}
    Respecting processor memories adds a constraint to a scheduling problem.
    Therefore, only specifically memory-targeted algorithms address this issue.
    Moreover, the way processor memories are represented in the model has a decisive impact on the way the constraint
    is formulated and addressed in the algorithm.
%
    Different models of memory available on processors and memory requirements of tasks have been presented.

    Marchal~et~al.~\cite{marchal2018parallel} assume a memory model where each processor has an individual memory available.
    Workflow tasks have no memory requirements, but they have input and output files that need to be stored in the memory.
    A polynomial-time algorithm for computing the peak memory needed for a parallel execution of such a DAG is provided,
    as well as an ILP solution to the scheduling problem.
    The memory model requires deleting all input data upon starting of the task and adding all output files there.

    In an assumed dual-memory systems~\cite{herrmann2014memory},  a processor can have access 
    to a memory of two different
    kinds (red or blue), and each task can be executed on only one sort of memory.
    The communications happen only between these two kinds of processors (communications within 
    each group are ignored).
    The authors then formulate an ILP solution for this problem formulation.

    %The algorithm presented by Yao et al.
    Yao et al.~\cite{yao2022memory} consider that each processor has an own internal memory and all
    processors share a common external one. The internal (local) memory is used to store the task files.
    The external memory is used to store evicted files to make room for the execution of a task on a processor.
    All processors, including the original one, can access these files. 
    Each edge %in~\cite{yao2022memory} 
    has two weights -- the size of the files transferred along it,
    and the time of communication along this edge.
    The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.

    In~\cite{ding2024ils}, there are connected processors with individual limited memories.
    The collective set of memories forms the global memory, to which each processor has access, however the access time
    to global memory is different.
    Each memory access in the graph is modeled as a memory access token on the task, while the edges have no weights.
    The solved problem is how to allocate initial input data in processor memories so that the overall
    execution is minimized and the memories are not exceeded.
    The authors propose an integer linear programming model.
    %that minimizes the length of the critical path, including a greedy initial solution.

    In~\cite{rodriguez2019exploration}, the authors assume memory requirement on tasks represented as tiles.
    Each processor has individual memories to process the task, but only the shared memories store the tiles containing
    memory tiles occupied by memory tiles.

   Finally, there are some cloud-oriented models that include costs associated with memory usage~\cite{liang2020memory}.

   Overall, there are a variety of memory models, but, to the best of our knowledge, the only study on a multiprocessor
   platform that is fully heterogeneous, with individual memories, is the one from~\cite{DBLP:conf/icpp/KulaginaMB24},
   but where a partition of the workflow is proposed, hence preventing processor reuse. 
   Hence, in ~\cite{DBLP:conf/icpp/KulaginaMB24}, there is no need of communication buffers to store data 
   that should be communicated
   between two processors when tasks are ready to execute.

    \subsection{Dynamic/adaptive algorithms}

We now review related work in a dynamic setting. With no variation in task parameters, 
    DVR HEFT~\cite{SANDOKJI2019482} rather considers that new tasks arrive in the system. 
    They use an almost unchanged HEFT algorithm in the static step, executing three slightly
    varying variants of task weighting and choosing the variant that gives the best overall makespan.
    In the dynamic phase, they receive new tasks and schedule them on either idle processors or 
    those processors that give them
    the earliest finish time.
    %Task failures are not covered.

    Rahman~\etal~\cite{rahman2013}'s dynamic critical path (DCP) algorithm for grids maps tasks to machines
    by calculating the critical path in the graph dynamically at every step.
    %For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
    %on the start time of a task (differing by the slack this task has).
    %All tasks on this critical path have the same earliest and latest start times, because they cannot be delayed.
    They schedule the first task on the critical path to the best suitable processor and recompute the critical path.
    %The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
    %If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum
    %execution time for the task is chosen.
    The heuristic also uses the same processor to schedule predecessor and successor tasks, as to avoid data transfer between processors.
    The approach is evaluated on random workflows of the size up to 300 tasks.

    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
    The procedure involves building a first (static) schedule, periodic resource monitoring and rescheduling the remaining
    tasks.
    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
    For each resource group, there is an own scheduler, and an overall global scheduler responsible for distributing
    tasks to groups.
    The static heuristic is HEFT with earliest start time as priority.
    Upon rescheduling, a new mapping is calculated from scratch, and this mapping is accepted if the resulting makespan
    is smaller than the previous one.
    The experiments were conducted on a single workflow with 10 tasks.
%
    % The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
    %The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.

    %The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
    %THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
    %They then go through tasks and map than to such processors that minimize the earliest start times of this task's
    %successors.
    %To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
    %average communication and computation costs fir the dependent tasks.
    %
    %The rescheduling is triggered when either a load on a resource increases over a threshold, or if a new resource
    %is added.

    Most dynamic or adaptive algorithms are formulated for clouds, where the execution environment is not fixed,
    but constrained by cost.

    Wang et al.~\cite{wang2019dynamic} propose a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
    Particles are possible solution in the solution space.
    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
    Similarly, Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.

    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling algorithm
    for clouds.
    They solve a set of linear equations that represent the cost of an execution based on the criteria.
    The authors test out 4 scenarios - one preferring each criteria, and a balanced one.
    The algorithm chooses the best virtual machine for each next task based on the cost given by the model.
    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
    %They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
 %   The authors use provenance data to make scheduling decisions.

    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
    on a single machine.
    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
    that these runtime changes are unrelated to each other).
    The authors formulate a decision space of all permutations of $n$ jobs, and the optimal schedule in relation to a
    performance measure $\phi$.
    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.

While several related work consider building a new schedule once some variation has been observed,
we are not aware of work implementing a real runtime system that interacts with the scheduler,
and tested on workflows with thousands of tasks, as we propose in this paper. Furthermore, 
we are not aware of any previous work discussing dynamic algorithms combined with memory constraints. 
%\AB{Say why our approach in dynamic setting is different and novel}

%    \subsection{Other notable works}
%
%    \cite{palis1996task} present a clustering-based scheduling algorithm for a parallel execution and prove its quality.
%    They utilize task duplication when creating the clusters (grains).
%    Their scheduler then maps clusters to processors.
%    They assume unlimited processors with the speed of 1.
%    For each task, they compute the earliest starting time and find a cluster, where this tasks's start time is as close
%    to it as possible.
%    %The cluster growing algorithm adds one task to the cluster at a time, by adding tasks in nondecreasing order of
%    %release times.
%
%    GRASP (generally randomized adaptive search procedure)~\cite{feo1989probabilistic} conducts a number of iterations
%    to search for an optimal solution for mapping tasks on machines.
%    A solution is generated at each step, and the best solution is kept at the end.
%    The search terminates when a certain termination criterion is reached.
%    It generates better results than other algorithms, because it explores the whole solution space.
%
%    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
%    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
%    The scenario contains AND- and OR-branches, where AND-branches indicate activities that need to be executed in parallel.
%    The heuristic first determines similar activities and groups them together.
%    Then they allocate these groups to disaster responders and tasks within this group according to a constraint system.
%    The dynamic part deals with changes and distinguishes between retriable and compensation activities.
%    The heuristic calculates a new execution path with these tasks.
%
%
%    \cite{lutke2024hetsim} is a scheduling simulator that models heterogeneous software with memory and accelerator
%    (processor) speed heterogeneity.
%    Each accelerator has its own memory that can be zero.
%    Each accelerator's characteristics depend on the task it runs and are not fixed.
%
%
%
%
%    \cite{meng2018traffic} investigate scheduling on multi-core chips.
%    Their model is far from ours.
%
%
%    An online scheduling algorithm~\cite{Witt2018POS} assumes a DAG-structured workflow and learns task characteristics.
%    They prioritize tasks that have failed before or are well-predictable.
%
%

\section{Model} %: \skug{Full: 4, polished: 3}}
\label{sec:model}

%\skug{
%CHANGES IN THIS CHAPTER
%}
    We first describe the model for our target applications, which are (large scientific) workflows for which we do not have perfect a priori knowledge,
    in Section~\ref{sec.mod.work}.  Next, we define the execution
    environment, a heterogeneous system (in terms of processor speed and memory size),
    in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in
    Section~\ref{sec.mod.pb}. The key notation is summarized in Table~\ref{tabnotation}.

    \subsection{Workflow}
    \label{sec.mod.work}
    A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
    $E$ is a set of directed edges of the form $e=(u,v)$, with \mbox{$u,v\in V$}, expressing precedence constraints between tasks.
    Each task~$u \in V$  is performing $w_u$ operations, and it also
    requires some amount of memory to be executed, denoted as~$m_u$.
    Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.

@@

%     the total memory requirement  for the execution of task~$u$ consists of the maximum
%    between the input files
%    (total size of the files to be received from the parents),
%    the output files (total size of the files to be sent to the children),
%    and the total memory size~$m_u$ (usually achieving the maximum):
%    \[
%        r_u = \max\left\{m_u , \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v}\right\}.
%    \]

@@

Furthermore, we place ourselves in a context where we do not have perfect knowledge
of the length of the tasks, i.e., $w_u$, % ($w_u$ and $m_u$) 
before the tasks complete their execution,
but only estimates~\cite{rahman2013,GARG2015256}.  
%. \AB{Add motivation: related work with variable task durations for instance...}
Hence, scheduling decisions are made on these estimated parameters, and
may be reconsidered at runtime when a task completes its execution. 
%when a task starts its execution and we know its exact parameters.
%\AB{Variability only on $w_u$ for now}

@@

%\skug{changes here}
The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
consists of $k$ processors $p_1, \dots, p_k$.
Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$ and a speed~$s_j$.
All processors have access to a shared disk of unlimited size, where they can write
and read data: Processor~$p_j$ has a bandwidth $\bw_j$ (resp.~$\br_j$) to write  (resp. read),
and hence it takes a time $\frac{c_{u,v}}{bw_j}$ to write the data produced by task~$u$
on~$p_j$ for task~$v$, 
if this data is evicted from memory.

%, but of much slower speed.
All communications happen over the disk -- after one processor has finished writing data there, another one can read
it to load the data on its individual memory. 
Hence, we can decide to evict some data from the individual processor memories if it has been 
written onto the disk, 
in order to free some space in the individual memories.
This data can later be read back into memory of the same or any other processor at any time.
%\skug {check this statement: If data has been written on disk, it is removed from the local memory (no two copies can be kept at the same time).} \AB{We can keep two copies: evict if we need to }

If the entire memory required by  task ~$u\in V$ fits in the available memory of processor~$p_j$,
    then the execution time of this task on this processor is expressed as $\frac{w_u}{s_j}$.
However, if the memory requirement of the task exceeds the available memory on the processor, 
then task can still be executed there, but it will be slowed down. Indeed, 
%    albeit slower.
the part of memory requirement that exceeds the available memory on processor~$j$ for task~$u$, 
denoted by $\Moff{u}$, %M\text{off}_u$, 
will be offloaded to the disk, slowing down the execution:

    \[
        w_{real} = \frac{w_u}{s_j} +  q_j \times  \Moff{u}, %  (m_u - Mav_u) , %   \frac{q_s \times w_u}{s_j} \frac{m_i - M_j}{m_i}
    \]
where $q_j$ is the slowdown coefficient for $p_j$. %, and $(m_u - Mav_u)$ is the part 
%of the task memory that needs to be offloaded.
\AB{Svetlana please check, I think we said in the end that we use absolute value 
of the amount of memory not fitting locally...}
%    \skug{check this equation}
%We assume that all processors are connected with the same bandwidth~$\beta$.
%    \hmey{Maybe mention that variable bandwidths are part of future work...?}

@@

We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
the memory requirement; it transforms the workflow into a series-parallel graph
and then finds the traversal that leads to the minimum memory consumption.
\AB{When do we need this? Probably more in Algorithms, right? }

%\skug{Todo: describe deviations}

>>>>

      TODO

@@

\section{Introduction}

    %%% CONTEXT %%%

    %%% MOTIVATION %%%

    %%% CONTRIBUTION %%%
    %\paragraph*{Contribution} 
    The main contributions of this paper are both algorithmic and experimental:
    \begin{itemize}
    \item We formalize the problem with memory constraints on heterogeneous processors, where a disk space
   stores files evicted from memory that will be later used by another processor.
    \item We design static and dynamic (adaptive?) heuristics
    \item We implement ...
    \item We perform extensive simulations,
    \item In the dynamic setting, we
    \end{itemize}

Todo: list what we do in each chapter

\section{Related work}
    \label{sec:related-work}

   Todo

\section{Model}
\label{sec:model}

\subsection{Workflow}
\label{sec.mod.work}

    Todo: this section holds. Needs rewriting to avoid plagiarism accusatons. Also insert other references from related work

Our workflow is a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
$E$ is a set of directed edges of the form $e=(u,v)$, with \mbox{$u,v\in V$}, expressing precedence constraints between tasks.
Each task~$u \in V$  is performing $w_u$ operations, and it also
requires some amount of memory to be executed, denoted as~$m_u$.
Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.

@@

@@

Furthermore, we place ourselves in a context where we do not have perfect knowledge
of the length of the tasks, i.e., $w_u$,
before the tasks complete their execution,
but only estimates~\cite{rahman2013,GARG2015256}.

Hence, scheduling decisions are made on these estimated parameters, and
may be reconsidered at runtime when a task completes its execution.

@@

    Todo: this section holds and is new (in parts).

The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
consists of $k$ processors $p_1, \dots, p_k$.
Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$ and a speed~$s_j$.
All processors have access to a shared disk of unlimited size, where they can write
and read data: Processor~$p_j$ has a bandwidth $\bw_j$ (resp.~$\br_j$) to write  (resp. read),
and hence it takes a time $\frac{c_{u,v}}{bw_j}$ to write the data produced by task~$u$
on~$p_j$ for task~$v$, 
if this data is evicted from memory.

All communications happen over the disk -- after one processor has finished writing data there, another one can read
it to load the data on its individual memory. 
Hence, we can decide to evict some data from the individual processor memories if it has been 
written onto the disk, 
in order to free some space in the individual memories.
This data can later be read back into memory of the same or any other processor at any time.

If the entire memory required by  task ~$u\in V$ fits in the available memory of processor~$p_j$,
    then the execution time of this task on this processor is expressed as $\frac{w_u}{s_j}$.
However, if the memory requirement of the task exceeds the available memory on the processor, 
then task can still be executed there, but it will be slowed down. Indeed,

the part of memory requirement that exceeds the available memory on processor~$j$ for task~$u$, 
denoted by $\Moff{u}$,
will be offloaded to the disk, slowing down the execution:

    \[
        w_{real} = \frac{w_u}{s_j} +  q_j \times  \Moff{u},
    \]
where $q_j$ is the slowdown coefficient for $p_j$.

@@

\textcolor{orange}{ TODO: Move the next paragraph to Algorithms.} We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
the memory requirement; it transforms the workflow into a series-parallel graph
and then finds the traversal that leads to the minimum memory consumption.
\AB{When do we need this? Probably more in Algorithms, right? }
